\Large{pipeComp, a general framework for the evaluation of computational pipelines, reveals performant single-cell RNA-seq preprocessing tools}

### 2020-03-07 main_figures/clustering.Rmd
updated figures to pdf

---
title: "Clustering"
author: "Pierre-Luc Germain"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    keep_md: true
---

@@

res <- readRDS("../data/results_clustering2.rds")
cl <- res$evaluation$clustering
cl$graph <- cl$clustmethod
levels(cl$graph) <- c(paste0("scran (",c("exact rank SNN", "Annoy rank SNN", "exact rank SNN", "exact KNN", "Annoy KNN", "exact SNN", "Annoy SNN"),")"), "Seurat SNN")
levels(cl$clustmethod) <- c("walktrap","walktrap","fast-greedy","walktrap","walktrap","walktrap","walktrap","Seurat")
cl$elapsed <- res$elapsed$stepwise$clustering$elapsed
cl$nbClust.diff <- cl$n_clus-cl$true.nbClusts
clo <- cl
cl <- cl[cl$k<=20,]
res$evaluation$clustering <- cl

renameFn <- function(x){
  x <- gsub("seurat; graph=|; graph=", "", gsub("seuratvst", "sctrans.+", x))
  x <- gsub("; clustmethod=", "+", x)
  x <- gsub("^graph=","",x)
  x <- gsub("exact ","",x)
  x <- gsub(" + walktrap","",x,fixed=TRUE)
}

@@

```{r clustering, fig.width=12, fig.height=8}
plot_grid(h2, h3, labels=LETTERS[1:2], ncol=1, rel_heights = c(7,4))
```
Clustering accuracy (A), computing time (B left) and mean deviation from the true number of clusters (B right). The (exact) rank-based SNN graph represents the default setting in `scran`-based clustering; however we find that the Annoy approximation to the KNN diminishes the computing costs without negatively affecting the accuracy.

>>>>

---
title: "Clustering"
author: "Pierre-Luc Germain"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

@@

res <- readRDS("../data/results_clustering.rds")
cl <- res$evaluation$clustering
cl$graph <- cl$clustmethod
levels(cl$graph) <- c(paste0("scran (",c("exact rank SNN", "Annoy rank SNN", "exact rank SNN", "exact KNN", "Annoy KNN", "exact SNN", "Annoy SNN"),")"), "Seurat SNN")
levels(cl$clustmethod) <- c("walktrap","walktrap","fast-greedy","walktrap","walktrap","walktrap","walktrap","Seurat")
cl$elapsed <- res$elapsed$stepwise$clustering$elapsed
cl$nbClust.diff <- cl$n_clus-cl$true.nbClusts
clo <- cl
cl <- cl[cl$k<=20,]
res$evaluation$clustering <- cl

renameFn <- function(x){
  x <- gsub("seurat; graph=|; graph=", "", gsub("seuratvst", "sctrans.+", x))
  x <- gsub("; clustmethod=", "+", x)
  x <- gsub("^graph=","",x)
  x <- gsub("exact ","",x)
  x <- gsub("+walktrap","",x,fixed=TRUE)
}

@@

```{r clustering, fig.width=12, fig.height=8}
pdf("clustering.pdf", width=15, height=10)
plot_grid(h2, h3, labels=LETTERS[1:2], ncol=1, rel_heights = c(7,4))
dev.off()
plot_grid(h2, h3, labels=LETTERS[1:2], ncol=1, rel_heights = c(7,4))
```
Clustering accuracy (A), computing time (B left) and mean deviation from the true number of clusters (B right). The (exact) rank-based SNN graph represents the default setting in `scran`-based clustering; however we find that the Annoy approximation to the KNN diminishes the computing costs without negatively affecting the accuracy.

### 2020-03-07 main_figures/dimensionality.Rmd
updated figures to pdf

---
title: "Dimensionality"
author: "Pierre-Luc Germain"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    keep_md: true
---

```{r}
suppressPackageStartupMessages({
  library(ggplot2)
  library(cowplot)
  library(ComplexHeatmap)
  library(matrixStats)
  library(pipeComp)
})
source("../misc_functions.R")
source(system.file("extdata/scrna_alternatives.R", package="pipeComp"))
theme_set(theme_cowplot(font_size = 12))
```

@@

```{r figure6, fig.width=9, fig.height=9.5}
plot_grid( p1, p2, p3, nrow=3, rel_heights=c(0.36,0.32,0.32), labels=c(NA,LETTERS[3:4]))
```

>>>>

---
title: "Dimensionality"
author: "Pierre-Luc Germain"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r}
suppressPackageStartupMessages({
  library(ggplot2)
  library(cowplot)
  library(ComplexHeatmap)
  library(matrixStats)
})
source("../misc_functions.R")
source(system.file("extdata/scrna_alternatives.R", package="pipeComp"))
theme_set(theme_cowplot(font_size = 12))
```

```{r, eval=FALSE}
# NB: the results were generated using an older version of pipeComp (branch 
# `biorxiv`), and must be plotted using this version of the package
suppressPackageStartupMessages({
  library(pipeComp)
})
```

```{r, include=FALSE}
#devtools::load_all("../../pipeComp_biorxiv/pipeComp/")
```

@@

```{r figure6, fig.width=9, fig.height=9.5}
pdf("dimensionality.pdf", width=9, height=10)
plot_grid( p1, p2, p3, nrow=3, rel_heights=c(0.36,0.32,0.32), labels=c(NA,LETTERS[3:4]))
dev.off()
plot_grid( p1, p2, p3, nrow=3, rel_heights=c(0.36,0.32,0.32), labels=c(NA,LETTERS[3:4]))
```

### 2020-03-07 main_figures/doublets.Rmd
updated figures to pdf

---
title: "Figure 1 doublets"
author: "Pierre-Luc Germain"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    keep_md: true

@@

```{r figure1, fig.height=6, fig.width=7}
p1m <- plot_grid(p1, p2, p3, nrow=1)
plot_grid(p1m, pt, p4, rel_heights=c(3.5,2,2), labels = LETTERS[1:3], ncol = 1, hjust=0)
```

>>>>

---
title: "Doublets figure"
author: "Pierre-Luc Germain"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document

@@

```{r figure1, fig.height=6, fig.width=7}
p1m <- plot_grid(p1, p2, p3, nrow=1)
pdf("doublets.pdf", width=6, height=7)
plot_grid(p1m, pt, p4, rel_heights=c(3.5,2,2), labels = LETTERS[1:3], ncol = 1, hjust=0, scale=0.95)
dev.off()
plot_grid(p1m, pt, p4, rel_heights=c(3.5,2,2), labels = LETTERS[1:3], ncol = 1, hjust=0)
```

### 2020-03-07 main_figures/dr.Rmd
updated figures to pdf

---
title: "Figure 5 dimensionality reduction"
author: "Pierre-Luc Germain"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    keep_md: true
---

```{r}
suppressPackageStartupMessages({
  library(ggplot2)
  library(cowplot)
  library(ComplexHeatmap)
  library(matrixStats)
  library(pipeComp)
})
source("../misc_functions.R")
```

@@

plot_grid(p1,p2, nrow=2, labels=LETTERS[1:2])
```

>>>>

---
title: "Figure 5 dimensionality reduction"
author: "Pierre-Luc Germain"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r}
suppressPackageStartupMessages({
  library(ggplot2)
  library(cowplot)
  library(ComplexHeatmap)
  library(matrixStats)
})
source("../misc_functions.R")
```

```{r, eval=FALSE}
# NB: the results were generated using an older version of pipeComp (branch 
# `biorxiv`), and must be plotted using this version of the package
suppressPackageStartupMessages({
  library(pipeComp)
})
```

```{r, include=FALSE}
#devtools::load_all("../../pipeComp_biorxiv/pipeComp/")
```

@@

pdf("dr.pdf", width = 12, height=6)
plot_grid(p1,p2, nrow=2, labels=LETTERS[1:2], scale=0.95)
dev.off()
plot_grid(p1,p2, nrow=2, labels=LETTERS[1:2])
```

### 2020-03-07 main_figures/filtering.Rmd
updated figures to pdf

---
title: "Figure 2 filtering"
author: "Pierre-Luc Germain"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    keep_md: true

---

```{r}
suppressPackageStartupMessages({
  library(ComplexHeatmap)
  library(circlize)
  library(ggplot2)
  library(ggridges)
  library(cowplot)
  library(SingleCellExperiment)
  library(pipeComp)
})
theme_set(theme_cowplot(font_size = 12))
source("../misc_functions.R")
data(datasets)
```

@@

```{r figure2, fig.width=8, fig.height=8.5}
res <- readRDS("../data/results_filtering.rds")
m <- scrna_evalPlot_filtering(res, returnTable=TRUE)
m$doubletRemoval <- !grepl("doubletmethod=none;", m$method)
m$filter <- factor(sapply(strsplit(as.character(m$method),"\\.|="), FUN=function(x) rev(x)[1]),
                     c("nofilter","lenient","default","pca","pca2","stringent","veryStringent"))
cols <- c("black",plgINS::getQualitativePalette(length(levels(m$filter))-1))
names(cols) <- levels(m$filter)

@@

>>>>

---
title: "Filtering figure"
author: "Pierre-Luc Germain"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r}
suppressPackageStartupMessages({
  library(ComplexHeatmap)
  library(circlize)
  library(ggplot2)
  library(ggridges)
  library(cowplot)
  library(SingleCellExperiment)
  library(pipeComp)
})
theme_set(theme_cowplot(font_size = 12))
source("../misc_functions.R")
```

@@

```{r figure2, fig.width=8, fig.height=8.5}
res <- readRDS("../data/results_filtering.rds")
m <- scrna_evalPlot_filtering(res, returnTable=TRUE)
m$doubletRemoval <- !grepl("doubletmethod=none;", m$method)
m$filter <- factor(sapply(strsplit(as.character(m$method),"\\.|="), FUN=function(x) rev(x)[1]),
                     c("nofilter","lenient","default","pca","pca2","stringent","veryStringent"))
cols <- c("black", pipeComp::getQualitativePalette(length(levels(m$filter))-1))
names(cols) <- levels(m$filter)

@@

pdf("filtering.pdf", width = 8, height=8.5)
plot_grid(p1, p2, nrow=2, rel_heights = c(3,8), labels=LETTERS[1:2])
dev.off()

### 2020-03-07 main_figures/imputation.Rmd
updated figures to pdf

---
title: "Imputation/denoising"
author: "Pierre-Luc Germain"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    keep_md: true
---

@@

```{r imputation, fig.width=10, fig.height=7}
plot_grid( p1, p2, nrow=2, labels=LETTERS[1:2], rel_heights = c(4,6))
```

>>>>

---
title: "Imputation/denoising"
author: "Pierre-Luc Germain"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

@@

```{r imputation, fig.width=10, fig.height=7}
pdf("imputation.pdf", width = 10, height=7)
plot_grid( p1, p2, nrow=2, labels=LETTERS[1:2], rel_heights = c(4.5,6))
dev.off()

plot_grid( p1, p2, nrow=2, labels=LETTERS[1:2], rel_heights = c(4,6))
```

### 2020-03-07 main_figures/selection.Rmd
updated figures to pdf

---
title: "Figure 4 selection"
author: "Pierre-Luc Germain"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    keep_md: true
---

```{r}
suppressPackageStartupMessages({
  library(SingleCellExperiment)
  library(ggplot2)
  library(cowplot)
  library(ComplexHeatmap)
  library(matrixStats)
  library(pipeComp)
})
source("../misc_functions.R")
data(datasets)
theme_set(theme_cowplot(font_size = 12))
```

@@

```{r figure4, fig.width=7.5, fig.height=6.5}
res <- readRDS("../data/results_selection.rds")
cl <- res$clustering
vst <- cl[grep("seuratvst",row.names(cl)),]
snorm <- cl[grep("seuratvst",row.names(cl), invert=TRUE),]
h1 <- scrna_evalPlot_clustAtTrueK(snorm[grep("=1000;",row.names(snorm)),], "ARI", scale = FALSE, title="ARI at true # clusters", anno_legend = FALSE)
h2 <- scrna_evalPlot_clustAtTrueK(vst[grep("=1000;",row.names(vst)),], "ARI", scale = FALSE, reorder_rows=row.names(h1@matrix), title="sctransform\nARI at true # clusters", name="ARI2", anno_legend = FALSE)
p2 <- grid.grabExpr(draw(h1+h2))
plot_grid(p1, p2, labels=LETTERS[1:2], nrow=2, rel_heights = c(4,5))
```

>>>>

---
title: "Selection figure"
author: "Pierre-Luc Germain"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r}
suppressPackageStartupMessages({
  library(SingleCellExperiment)
  library(ggplot2)
  library(cowplot)
  library(ComplexHeatmap)
  library(matrixStats)
})
source("../misc_functions.R")
data(datasets)
theme_set(theme_cowplot(font_size = 12))
```

```{r, eval=FALSE}
# NB: the results were generated using an older version of pipeComp (branch 
# `biorxiv`), and must be plotted using this version of the package
suppressPackageStartupMessages({
  library(pipeComp)
})
```

```{r, include=FALSE}
#devtools::load_all("../../pipeComp_biorxiv/pipeComp/")
```

@@

```{r}
res <- readRDS("../data/results_selection.rds")
cl <- res$clustering
vst <- cl[grep("seuratvst",row.names(cl)),]
snorm <- cl[grep("seuratvst",row.names(cl), invert=TRUE),]
h1 <- scrna_evalPlot_clustAtTrueK(snorm[grep("=1000;",row.names(snorm)),], "ARI", scale = FALSE, title="ARI at true # clusters", anno_legend = FALSE)
h2 <- scrna_evalPlot_clustAtTrueK(vst[grep("=1000;",row.names(vst)),], "ARI", scale = FALSE, reorder_rows=row.names(h1@matrix), title="sctransform\nARI at true # clusters", name="ARI2", anno_legend = FALSE)
p2 <- grid.grabExpr(draw(h1+h2))
```

```{r figure4, fig.width=7.5, fig.height=6.5}
pdf("selection.pdf", width = 7.5, height=6.5)
plot_grid(p1, p2, labels=LETTERS[1:2], nrow=2, rel_heights = c(4,5))
dev.off()

plot_grid(p1, p2, labels=LETTERS[1:2], nrow=2, rel_heights = c(4,5))
```

### 2020-03-07 supplementary_figures/supp1b_metrics.Rmd
updated figures to pdf

```{r, include=FALSE}
res <- readRDS("../data/results_clustering.rds")
res <- res$evaluation$clustering
res <- res[which(res$clustmethod %in% c("clust.seurat","clust.scran")),]
res$norm <- gsub("norm\\.seuratvst","sctransform",res$norm)
dd <- res[!duplicated(res[,c("dataset","true.nbClusts")]),]
```

```{r, fig.width=7.5, fig.height=7}
ggplot(res, aes(n_clus, ARI, colour=clustmethod, shape=norm)) + 
    geom_point() + xlim(c(1,40)) + facet_wrap(~dataset) + xlab("Number of clusters") +
    geom_vline(aes(xintercept=k), data=dd, linetype="dashed") + theme(legend.position="bottom") + 
    guides(colour=guide_legend(ncol=1, title = ""), shape=guide_legend(ncol=1, title=""))
```

>>>>

```{r, include=FALSE}
res <- readRDS("../data/results_clustering.rds")
res <- res$evaluation$clustering
res <- res[which(res$clustmethod %in% c("clust.seurat","clust.scran")),]
res$norm <- gsub("norm\\.seuratvst","sctransform",res$norm)
dd <- res[!duplicated(res[,c("dataset","true.nbClusts")]),]
p <- ggplot(res, aes(n_clus, ARI, colour=clustmethod, shape=norm)) + 
    geom_point() + xlim(c(1,40)) + facet_wrap(~dataset) + xlab("Number of clusters") +
    geom_vline(aes(xintercept=k), data=dd, linetype="dashed") + theme(legend.position="bottom") + 
    guides(colour=guide_legend(ncol=1, title = ""), shape=guide_legend(ncol=1, title=""))
```

```{r, fig.width=7.5, fig.height=7}
p
```

### 2020-03-07 supplementary_figures/supp_figures.Rmd
updated supplementary figures
updated figures to pdf

---
title: "Supplementary Figures"
author: "Pierre-Luc Germain"
date: "`r format(Sys.time(), '%d %B, %Y')`"
geometry: margin=1.5cm
output:
  pdf_document:
    keep_tex: true
    toc: true
    fig_width: 8
    fig_height: 6
    citation_package: natbib
bibliography: ../bmc_article.bib
---

@@

>>>>

---
title: \Large{pipeComp, a general framework for the evaluation of computational pipelines, reveals performant single-cell RNA-seq preprocessing tools}
subtitle: \vspace{0.2cm}\LARGE{Supplementary Figures}\vspace{0.3cm}
author:
- "Pierre-Luc Germain"
- "Anthony Sonrel"
- "Mark D. Robinson"
geometry: margin=1.5cm
output:
  pdf_document:
    toc: true
    fig_width: 8
    fig_height: 6
    citation_package: natbib
bibliography: ../bmc_article.bib
---

\renewcommand{\thepage}{S\arabic{page}}

@@

```{r child='supp1b_metrics.Rmd'}
```
